---
title: "Git, GitHub & RStudio"
description: RStudio users can leverage git Version control to enhance reproducibility.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

```

The [`usethis`](https://usethis.r-lib.org/) package and the RStudio IDE will help you orchestrate your version control repositories as RStudio projects.  

<iframe width="560" height="315" src="https://www.youtube.com/embed/r5_v5f6128M" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<!-- <iframe width="560" height="315" src="https://warpwire.duke.edu/w/t0oFAA/" frameborder="0" scrolling="0" allow="autoplay; encrypted-media; fullscreen;  picture-in-picture;" allowfullscreen></iframe> -->

<iframe width="560" height="315" src="https://www.youtube.com/embed/Cn-72tbRNFc" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


### Setup

To get started, make a free GitHub.com account.  Make sure you have the latest versions of R, RStudio, and usethis.  Then, setup your RStudio configuration using two key documents:  [Setup](https://usethis.r-lib.org/articles/articles/usethis-setup.html), and [Managing Git(Hub) Credentials](https://usethis.r-lib.org/articles/articles/git-credentials.html).

### Tips

#### Clone or fork from GitHub

```{r echo=TRUE, eval=FALSE}
usethis::create_from_github("https://github.com/data-and-visualization/git-tutorial")
```

#### Create the GitHub repo from local

```{r echo=TRUE, eval=FALSE}
library(usethis)
use_git()          # initialize a git repo
use_git_ignore()
use_readme.Rmd()
use_ccby_license()
# Before next step  → stage & commit changes
use_github()       # Connect local repo to GitHub
```

#### Rename default branch

```{r echo=TRUE, eval=FALSE}
library(usethis)
git_default_branch()
git_default_branch_rename()
git_default_branch_rediscover()
```

#### set-up git and manage GitHub credentials

```{r echo=TRUE, eval=FALSE}
# Sign-in to your GitHub account
# Read: https://usethis.r-lib.org/articles/articles/git-credentials.html
# Read: https://usethis.r-lib.org/articles/articles/usethis-setup.html
usethis::create_github_token()
gitcreds::gitcreds_set()
gh::gh_whoami()
usethis::git_sitrep()
usethis::use_git_config(user.name = "Jane Doe", user.email = "jane@example.com")
usethis::git_sitrep()
```

### Next Level...

Take your reproducibility to the next level.  [Make your code citable](https://guides.github.com/activities/citable-code/) by connecting your GitHub repo and your [ORCID](https://orcid.org/register) (unique author ID) through to the [Zenodo](https://zenodo.org/) archival repository.  Now, every time you commit a milestone release, you also mint a DOI.  

#### Share a zero-install compute environment

<!-- Containerize your repository with binder.org using the [`holepunch` package](https://karthik.github.io/holepunch/) -->

Publish your code in a binder.org container

<!-- Use the [`holepunch` package](https://karthik.github.io/holepunch/) -->

```{r echo=FALSE, eval=FALSE}
library(holepunch)

write_compendium_description()
write_dockerfile()
generate_badge()
build_binder()

```

<!-- #### Bespoke_Publish your code in a binder  -->

<!-- I recently found my QuickPublish approach _didn't quite work_.  So I RTFMed the binder.org docs and it seemed to imply a preference for NOT using a `.binder/Dockerfile`.  Therefore, here is an approach that is slightly less automated but more bulletproof. -->

Everything you need is exemplified in binder-example for R that you can [find on github](https://github.com/binder-examples/r)

The key elements are an `install.R` and `runtime.txt` files.  The content and format of each file is demonstrated properly at the GitHub repo.  Browse and you will see.

**Step 1**

```{r echo=FALSE, eval=FALSE}
# holepunch::write_compendium_description()  # output is a DESCRIPTION file in the project root
holepunch::write_runtime()    # output is .binder/runtime.txt
holepunch::write_install()    # output is .binder/install.R
holepunch::generate_badge()   # output is a badge to be found in the README file(s)
```

a. Create an `install.R` file and store it in the RStudio Project's root directory.<br> 
Follow the example at https://github.com/binder-examples/r <br>

> This file should consist of install functions for each package used in your R code.  If you're not sure what packages you're using, try running `holepunch::write_compendium_description()` and you will find the details listed in the DESCRIPTION file under the 'Depends' and 'Remotes' sections. **IMPORTANT**:  Delete the `DESCRIPTION` file Before launching the binder build.  In my experience the DESCRIPTION file will conflict with the binder's build process. <br>
For non-CRAN packages -- specifically packages installed from GitHub -- I recommend using the format: `remotes::install_github("gadenbuie/countdown")` for each package that is not on CRAN.  CRAN packages can follow the format: `install.packages("dplyr")`

b. Create a `runtime.txt` file and store it in the RStudio Project's root directory<br>
Follow the example at https://github.com/binder-examples/r <br>

> Note:  If you try using `holepunch`, it will store these files in a `.binder` subdirectory.  In my experience this will cause the binder build to fail.  For this reason, I don't recommend using holepunch and I do recommend storing the two files in the project's root directory.

**Step 2**

Generate a binder badge with `usethis::use_binder_badge(urlpath = "rstudio")`.  This should store the badge in your GitHub's project `README` file.  Then `commit` and `push` your changes to GitHub; go to the README on GitHub, launch the build and wait for the long launch/build time.

Hint:  be prepared to wait.  The first build can take some time.  Sometimes you must launch an initial or dormant build more than once.  In my experienced these hiccups are rare, but try the launch twice before getting involved in extensive troubleshooting.

<!-- a. Edit the `.binder/install.R` file using the following [model](https://github.com/binder-examples/r/blob/master/install.R). -->
<!-- b. Edit the `binder/install.R` file based on this [model](https://github.com/binder-examples/r/blob/master/runtime.txt) or the [documentation](https://github.com/jupyterhub/repo2docker/pull/772#issue-313426641) -->
<!-- b. Edit the `DESCRIPTION` file using this basic [model](https://github.com/libjohn/workshop_twitter_analysis/blob/main/DESCRIPTION). **Modify all the fields**.  You may want to base the `Depends` and `Remotes` sections on your `.binder/install.R` file -->

---

_The old documentation from this site can be found in the [v.0.2019 release](https://github.com/data-and-visualization/git-tutorial/releases/tag/v0.9.2019) on GitHub_. 
