---
title: "Git, GitHub & RStudio"
description: RStudio users can leverage git Version control to enhance reproducibility.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

```

The [`usethis`](https://usethis.r-lib.org/) package and the RStudio IDE will help you orchestrate your version control repositories as RStudio projects.  

<iframe width="560" height="315" src="https://www.youtube.com/embed/r5_v5f6128M" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

<!-- <iframe width="560" height="315" src="https://warpwire.duke.edu/w/t0oFAA/" frameborder="0" scrolling="0" allow="autoplay; encrypted-media; fullscreen;  picture-in-picture;" allowfullscreen></iframe> -->

<iframe width="560" height="315" src="https://www.youtube.com/embed/Cn-72tbRNFc" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>



### Setup

To get started, make a free GitHub.com account.  Make sure you have the latest versions of R, RStudio, and usethis.  Then, setup your RStudio configuration using two key documents:  [Setup](https://usethis.r-lib.org/articles/articles/usethis-setup.html), and [Managing Git(Hub) Credentials](https://usethis.r-lib.org/articles/articles/git-credentials.html).

### Tips

#### Clone or fork from GitHub

```{r echo=TRUE, eval=FALSE}
usethis::create_from_github("https://github.com/data-and-visualization/git-tutorial")
```

#### Create the GitHub repo from local

```{r echo=TRUE, eval=FALSE}
library(usethis)
use_git()          # initialize a git repo
use_git_ignore()
use_readme.Rmd()
use_ccby_license()
use_github()       # Connect local repo to GitHub
```

#### set-up git and manage GitHub credentials

```{r echo=TRUE, eval=FALSE}
# Sign-in to your GitHub account
# Read: https://usethis.r-lib.org/articles/articles/git-credentials.html
# Read: https://usethis.r-lib.org/articles/articles/usethis-setup.html
usethis::create_github_token()
gitcreds::gitcreds_set()
gh::gh_whoami()
usethis::git_sitrep()
usethis::use_git_config(user.name = "Jane Doe", user.email = "jane@example.com")
usethis::git_sitrep()
```

### Next Level...

Take your reproducibility to the next level.  [Make your code citable](https://guides.github.com/activities/citable-code/) by connecting your GitHub repo and your [ORCID](https://orcid.org/register) (unique author ID) through to the [Zenodo](https://zenodo.org/) archival repository.  Now, every time you commit a milestone release, you also mint a DOI.  

Containerize your repository with binder.org using the [`holepunch` package](https://karthik.github.io/holepunch/).

#### Quick_Publish your code in a binder.org container

Use the [`holepunch` package](https://karthik.github.io/holepunch/)

```{r echo=TRUE, eval=FALSE}
library(holepunch)

write_compendium_description()
write_dockerfile()
generate_badge()
build_binder()

```

#### Bespoke_Publish your code in a binder 

I recently found my QuickPublish approach _didn't quite work_.  So I RTFMed the binder.org docs and it seemed to imply a preference for NOT using a `.binder/Dockerfile`.  Therefore, here is an approach that is slightly less automated but more bulletproof.

**Step 1**

```{r echo=TRUE, eval=FALSE}
holepunch::write_compendium_description()  # output is a DESCRIPTION file in the project root
holepunch::write_runtime()    # output is .binder/runtime.txt
holepunch::write_install()    # output is .binder/install.R
holepunch::generate_badge()   # output is a badge to be found in the README file(s)
```

**Step 2**


a. Edit the `.binder/install.R` file using the following [model](https://github.com/binder-examples/r/blob/master/install.R).
b. Edit the `binder/install.R` file based on this [model](https://github.com/binder-examples/r/blob/master/runtime.txt) or the [documentation](https://github.com/jupyterhub/repo2docker/pull/772#issue-313426641)
b. Edit the `DESCRIPTION` file using this basic [model](https://github.com/libjohn/workshop_twitter_analysis/blob/main/DESCRIPTION). **Modify all the fields**.  You may want to base the `Depends` and `Remotes` sections on your `.binder/install.R` file

***
test test

---

_The old documentation from this site can be found in the [v.0.2019 release](https://github.com/data-and-visualization/git-tutorial/releases/tag/v0.9.2019) on GitHub_. 
