---
title: "Git, GitHub & RStudio"
description: Makin' it work!!
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

```

The [`usethis`](https://usethis.r-lib.org/) package and the RStudio IDE will help you orchestrate your version control repositories as RStudio projects.  

<iframe width="560" height="315" src="https://www.youtube.com/embed/p8q4WGg6how" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

### Setup

To get started, make a free GitHub.com account.  Make sure you have the latest versions of R, RStudio, and usethis.  Then, setup your RStudio configuration using two key documents:  [Setup](https://usethis.r-lib.org/articles/articles/usethis-setup.html), and [Managing Git(Hub) Credentials](https://usethis.r-lib.org/articles/articles/git-credentials.html).

### Tips

#### Clone or fork from GitHub

```{r echo=TRUE, eval=FALSE}
usethis::create_from_github("https://github.com/data-and-visualization/git-tutorial")
```

#### Create the GitHub repo from local

```{r echo=TRUE, eval=FALSE}
library(usethis)
use_git()          # initialize a git repo
use_git_ignore()
use_readme.Rmd()
use_ccby_license()
use_github()       # Connect local repo to GitHub
```

#### set-up git and manage GitHub credentials

```{r echo=TRUE, eval=FALSE}
# Sign-in to your GitHub account
# Read: https://usethis.r-lib.org/articles/articles/git-credentials.html
# Read: https://usethis.r-lib.org/articles/articles/usethis-setup.html
usethis::create_github_token()
gitcreds::gitcreds_set()
gh::gh_whoami()
usethis::git_sitrep()
usethis::use_git_config(user.name = "Jane Doe", user.email = "jane@example.com")
gh::git_sitrep()
```

### Next Level...

Take your reproducibility to the next level.  [Make your code citable](https://guides.github.com/activities/citable-code/) by connecting your GitHub repo and your [ORCID](https://orcid.org/register) (unique author ID) through to the [Zenodo](https://zenodo.org/) archival repository.  Now, every time you commit a milestone release, you also mint a DOI.  

Containerize your repository with binder.org using the [`holepunch` package](https://karthik.github.io/holepunch/).

#### Publish your code in a binder.org container

Use the [`holepunch` package](https://karthik.github.io/holepunch/)

```{r echo=TRUE, eval=FALSE}
library(holepunch)

write_compendium_description()
write_dockerfile()
generate_badge()
build_binder()

```

---

_The old documentation from this site can be found in the [v.0.2019 release](https://github.com/data-and-visualization/git-tutorial/releases/tag/v0.9.2019) on GitHub_. 
